---
title: "Machine Learning - Course WriteUp"
author: "Zacarías Ojeda"
date: "August 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
set.seed(123)
```

# Practical Machine Learning - Course Project

*Prediction Assignment Writeup*


## Load data

```{R}
data <- read.csv('pml-training.csv',na.strings = c("NA", "#DIV/0!"))
```

## Exploratory Data Analysis

```{R}
dim(data)
```

We have a lot of variables (160), lets try to reduce this dimension

```{R}
nZero <- nearZeroVar(data, freqCut = 95/5, uniqueCut = 10, saveMetrics = T)
data <- data[,nZero$nzv == FALSE]

dim(data)
```

Still a large number of variables, lets try to remove those columns with high percentage of NA values (90%).

```{R}
data <- data[, -which(colMeans(is.na(data)) > 0.9)]

dim(data)
```

A lot of variables were removed, lets try to inspect for not useful variables for data purposes.

```{R}
names(data)
``` 

The index value (X), user_name and time related are not interesting to traint our prediction models, 
```{R}
data <- data[, -c(1,2,3,4,5)]

dim(data)
```


## Creating a prediction model

### Data slicing ()

```{r}
inTrain <- createDataPartition(y=data$classe, p=0.7, list=FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]
  
```
### Training a Generalized Boosted Regression predicion model

Using K-fold cross validation
```{r}
train_control <- trainControl(method="cv", number=3)

```

```{r}
system.time(model_gbm <- train(classe ~ .,method="gbm",data=training,trControl=train_control, verbose=F))
```
```{r}
plot(model_gbm)
```


#### Estimating Accuracy

```{r}
confusionMatrix(testing$classe,predict(model_gbm,testing))
```

### Training a RandomForest predicion model

```{r}
system.time(model_rf <- train(classe ~ .,method="rf",data=training,trControl=train_control, verbose=F))
```

```{r}
plot(model_rf)
```


#### Estiamating Accuracy

```{r}
confusionMatrix(testing$classe,predict(model_rf,testing))
```


It seems that Random Forest performs better.

## Predicting Values using 'pml-testing.csv' dataset


```{r}
to_predict <- read.csv('pml-testing.csv',na.strings = c("NA", "#DIV/0!"))
```

```{r}
predictions_rf <- predict(model_rf, newdata = to_predict)
predictions_rf
```

